\documentclass[11pt]{article}

\usepackage{pstricks,pst-node,pst-tree}
\usepackage[noend]{algpseudocode}

\pagestyle{empty}
\textwidth	165mm
\textheight	252mm
\topmargin	-13mm
\oddsidemargin	-2mm
\evensidemargin	-2mm
\renewcommand{\baselinestretch}{1.03}

\newcommand{\id}[1]{\mbox{\textit{#1}}}
\newcommand{\tuple}[1]{\langle #1 \rangle}

\algnewcommand{\AND}{\,\textbf{and}\,}
\algnewcommand{\NOT}{\,\textbf{not}\,}
\algnewcommand{\OR}{\,\textbf{or}\,}
\algnewcommand{\TRUE}{\,\texttt{true}\,}
\algnewcommand{\FALSE}{\,\texttt{false}\,}
\MakeRobust{\Call}
\begin{document}

\begin{center}
{\sc The University of Melbourne
\\
School of Computing and Information Systems
\\
COMP90038 Algorithms and Complexity}
\bigskip \\
{\Large\bf Assignment 2, Semester 2, 2017}
\bigskip \\
{\large Released: 23 September.  Deadline: 13 October at 23:00}
\end{center}

\section*{Objectives}

To improve your understanding of data structures and algorithms for
sorting and search.
To consolidate your knowledge of trees and tree-based algorithms.
To develop problem-solving and design skills.
To develop skills in analysis and formal reasoning about
complex concepts.
To improve written communication skills; in particular the
ability to use pseudo-code and present algorithms clearly,
precisely and unambiguously.

\section*{Five Challenges}

\subsection*{Challenge 1 \hfill {\small (1 mark)}}

Draw a valid AVL tree of height 5, consisting of the 20 integers
from 1 to 20 (no repeats).
Note that a tree of height 5 has some root-to-leaf path with
\emph{six} nodes along it.
\\
\textsc{answer}
\begin{center}
\pstree[nodesep=2pt,levelsep=25pt]{\TR{13}}{
  \pstree[nodesep=2pt,levelsep=25pt]{\TR{8}}{
    \pstree{\TR{5}}{
      \pstree{\TR{3}}{
        \pstree{\TR{2}}{
          \TR{1}
        }
        \TR{4}
      }
      \pstree{\TR{7}}{
        \TR{6}
      }
    }
    \pstree{\TR{11}}{
      \pstree{\TR{10}}{
        \TR{9}
      }
      \TR{12}
    }
  }
  \pstree[nodesep=2pt,levelsep=25pt]{\TR{18}}{
    \pstree{\TR{16}}{
      \pstree{\TR{15}}{
        \TR{14}
      }
      \TR{17}
    }
    \pstree{\TR{20}}{
      \TR{19}
    }
  }
}
\end{center}
For a AVL tree which contains 20 integers, it has the minimum height of 4
when it is a complete binary tree. To find the maxium height,
we try to find the minimum nodes number given a specific height.
So, if $n(h)$ means the minimum nodes number of height h, then
$$n(h) = n(h-1) + n(h-2) + 1$$
since we want to find the minimum nodes, so we want the heights of subtrees
are as small as possible.
Then we can find $n(5) = 20$, which means, to achieve height of 5,
we need 20 nodes at least.
So, 5 is the maxium height of 20 nodes AVL tree.
To construct the tree, we obey the rule $n(h) = n(h-1) + n(h-2) + 1$,
and after we get the structure,
we could use inorder traverse to fit the integers into the structure.
\pagebreak

\subsection*{Challenge 2 \hfill {\small (2 marks)}}

Snarkbusters Inc.\ produces and deploys snark detectors,
to be used on building facades.
These are very expensive devices.
One of the company's workers dropped a snark detector while working on the
16th floor of a building in Melbourne's CBD, and the device broke when it
hit the ground, in spite of having a reputation as indestructible.
The company now wants to know how tough their snark detector really is.
They have hired a consultant, Dr.\ Eva Luator, to determine the
largest floor level from which a snark detector can be dropped
without breaking.
That is, they want the number $n$ such that dropping a snark detector
from level $n$ is safe (it does not break), but dropping it from
level $n+1$ is unsafe (it will break).
They call that number $n$ the ``safe limit''.
They know that there is no problem when the device is dropped from ground
level, that is, they know that $0 < n < 16$.

Since there are 15 floor levels to test, Dr.\ Luator asks for a
handful of snark detectors to use in the experiment.
Once a snark detector is broken it cannot be used again in the test.
But the company will not give her that many.
They argue that one is enough, because Dr.\ Luator can just drop it from
level 1, and then, if it did not break, drop it from level 2, and so on.
That process will identify the safe limit.
Dr.\ Luator protests: that could make the testing too time consuming
(and expensive) because, in the worst case, she would need to perform
15 experiments (drops).
As a compromise, it is decided that Dr.\ Luator can have two snark
detectors to use in her testing.

Eva Luator wants to make the most of the two snark detectors.
She wants to minimise the number $d$ of experiments (drops) that she
needs to do, \emph{in the worst case}, to determine the safe limit.
What is that number $d$, and what is the testing strategy that achieves it?
\\
\textsc{answer}
\par

$d = 5$

To solve this problem, we could construct a model, which contains the ordered
integers
from 1 to 15, and a given float m,
and our goal is to fit m into the ordered intergers,
but we only have two chance to compare m with a bigger integer.

The most straightfoward idea is that we use binary search,
and once meet the bigger integer,
then we do a linear search on the small part from the lowest integer to the highest integer.

However, this will cost 8 comparisons in the worst case.
The optimization is to balance the iterations of linear search and binary search.
To achieve that, we construct a binary search tree shown as below.
The tree is optimized since the number of nodes in the left subtree is equal to
the height of right subtree, which give a minimum time of comparisons, 5.
\begin{center}
\pstree[nodesep=2pt,levelsep=25pt]{\TR{5}}{
  \pstree[nodesep=2pt,levelsep=25pt]{\TR{3}}{
    \pstree{\TR{2}}{
      \TR{1}
    }
    \TR{4}
  }
  \pstree[nodesep=2pt,levelsep=25pt]{\TR{9}}{
    \pstree{\TR{7}}{
      \TR{6}
      \TR{8}
    }
    \pstree{\TR{12}}{
      \pstree{\TR{11}}{
        \TR{10}
      }
      \pstree{\TR{14}}{
        \TR{13}
        \TR{15}
      }
    }
  }
}
\end{center}


\pagebreak
\subsection*{Challenge 3 \hfill {\small (3 marks)}}

Consider the following problem.
We are given two arrays, $A[1]..A[n]$ and $B[1]..B[n]$,
each holding positive integer keys (possibly with duplicates).
The arrays are of the same size, and each is sorted.
We are also given a positive integer $s$, and we want to find a
pair $(i,j)$ of indices such that $A[i] + B[j] = s$.
If several such pairs exist, we do not care which is produced---any
such pair will do.
If no such pair exists then we want that indicated,
in the form of $(0,0)$, a non-valid pair of indices, being returned.
Using neat, clean pseudo-code, write three different functions to
solve the problem:
\begin{enumerate}
\setlength{\itemsep}{-0.4ex}
\item
A brute-force function $\textsc{F1}(A[\cdot],B[\cdot],n,s)$.
The worst-case running time of \textsc{F1} should be $\Theta(n^2)$.
Make \textsc{F1} as simple and neat as you can.
\\
\textsc{answer}
\par
\begin{algorithmic}
\Function{F1}{$A[\cdot],B[\cdot],n,s$}
\For{$i \gets 1,n$}
  \For{$j \gets 1, n$}
    \If{$A[i]+B[j]=s$}
      \State \Return $(i, j)$
    \EndIf
  \EndFor
\EndFor
\State \Return $(0, 0)$
\EndFunction
\end{algorithmic}
\item
A function
$\textsc{F2}(A[\cdot],B[\cdot],n,s)$.
The worst-case running time of \textsc{F2} should be $\Theta(n \log n)$.
Make \textsc{F2} as simple and neat as you can.
\\
\textsc{answer}
\par
\begin{algorithmic}
  \Function{F2}{$A[\cdot],B[\cdot],n,s$}
    \For{$i \gets 1,n$}
      \State $low \gets 1$
      \State $high \gets n$
    \While{$low < high$}
      \State $m \gets floor((high - low)/2) + low$
      \If{$A[i]+B[m]=s$}
        \State \Return $(i, m)$
      \ElsIf{$A[i]+B[m]<s$}
        \State{$low \gets m+1$}
      \Else
        \State{$high \gets m-1$}
      \EndIf
    \EndWhile
    \EndFor
  \State \Return $(0,0)$
  \EndFunction

\end{algorithmic}
\item
A function $\textsc{F3}(A[\cdot],B[\cdot],n,s)$.
The worst-case running time of \textsc{F3} should be $\Theta(n)$.
Make \textsc{F3} as simple and neat as you can.
\\
\textsc{answer}
\par
\begin{algorithmic}
  \Function{F3}{$A[\cdot],B[\cdot],n,s$}
    \State $left \gets 1$
    \State $right \gets n$
    \While{$left \leq n \AND right \geq 1$}
    \If{$A[left]+B[right]=s$}
      \State \Return $(left, right)$
    \ElsIf{$A[i]+B[m]<s$}
      \State{$left \gets left+1$}
    \Else
      \State{$right \gets right-1$}
    \EndIf
    \EndWhile
  \State \Return $(0,0)$
  \EndFunction
\end{algorithmic}
\end{enumerate}
\pagebreak

\subsection*{Challenge 4 \hfill {\small (2 marks)}}

A $d$-dimensional vector $\tuple{u_1,u_2,\ldots,u_d}$ \emph{screens}
the vector $\tuple{v_1,v_2,\ldots,v_d}$ iff there is a permutation
$\pi$ on $\{1,2,\ldots,d\}$ such that
$u_1 > v_{\pi(1)}$,
$u_2 > v_{\pi(2)}$,
\ldots,
$u_d > v_{\pi(d)}$.
For example, $\tuple{7,9,4}$ screens $\tuple{7,6,3}$
because one permutation of the latter is $\tuple{6,7,3}$.
On the other hand, $\tuple{7,9,4,5}$ does not screen $\tuple{7,6,3,5}$.
The last example also shows that we can have vectors
$\mathbf{u}$ and $\mathbf{v}$, neither of which screens the other.

Using pseudo-code,
give an algorithm which determines, given two vectors $\mathbf{u}$
and $\mathbf{v}$, whether $\mathbf{u}$ screens $\mathbf{v}$.
It should return True iff $\mathbf{u}$ screens $\mathbf{v}$.
If it helps, you an assume that vectors are arrays.
Make the algorithm as efficient as you can, and state,
as precisely as you can, its worst-case time complexity
(and, if possible, average-case as well).
\\
\textsc{answer}
\par
\begin{algorithmic}
  \Function{doesScreen}{$u[\cdot], v[\cdot], d$}
    \State buildHeap($u[\cdot], d$)\Comment{build the max heap}
    \State buildHeap($v[\cdot], d$)
    \For{$i \gets 1,d$}
      \If{$ \Call{eject}{u[\cdot], d-i+1} \leq \Call{eject}{v[\cdot], d-i+1}$}
        \State \Return $\FALSE$
      \EndIf
    \EndFor
    \State \Return $\TRUE$
  \EndFunction

  \Function{buildHeap}{$A[\cdot], n$}
    \For{$i \gets floor(n/2)$ downto $1$}
      \State $\Call{restoreHeap}{A[\cdot], i, n}$
    \EndFor
  \EndFunction

  \Function{eject}{$A[\cdot], n$}
    \State $item \gets A[1]$
    \State $A[1] \gets A[n]$
    \State $A[n] \gets item$
    \State $\Call{restoreHeap}{A[\cdot], 1, n-1}$
    \State \Return $item$
  \EndFunction

  \Function{restoreHeap}{$A[\cdot], k, n$}\Comment{restore a heap at parent node k}
    \State $v \gets A[k]$
    \State $heap \gets \FALSE$
    \While{$\NOT heap \AND 2 * k \leq n$}
      \State $j \gets 2 * k$
      \If{$j < n \AND A[j] < A[j+1]$}
        \State $j \gets j+1$
      \EndIf
      \If{$v \geq A[j]$}
        \State $heap \gets \TRUE$
      \Else
        \State $A[k] \gets A[j]$
        \State $k \gets j$
      \EndIf
    \EndWhile
    \State $A[k] \gets v$
  \EndFunction

\end{algorithmic}

To judge whether the vector $u$ screens the $v$,
it could be the same problem of sorting two arrays(treat vector as array),
and judge whether every element in one array are bigger than the corresponding
element in the other array.

I use heap sort to achieve this goal. The worst case and average complexity are
both $\Theta(n \log n)$,
since building heap is $\Theta(n)$ and ejecting is $\Theta(n \log n)$
in the worst case.
However, by using the heap, we do not need to sort the
entire array, but restore it to a heap every time after ejecting, so the pratical
performance could be better than other sorting algorithms.
\pagebreak
\subsection*{Challenge 5 \hfill {\small (2 marks)}}

An array-based implementation of priority queues is not ideal if we
often need to ``merge'' several priority queues into one, or if we
are programming in a language that does not have arrays.
Here we are interested in a data structure that is \emph{adaptive}
and supports efficient merging, by which we mean taking two
priority queues $P$ and $P'$ and forming a new priority queue with
elements $P \cup P'$.
As is often the case with adaptive data structures, some operations
may be worst-case expensive, but when we average their cost across many
applications, they are cheap.
That will be the case for the operations we want to implement here:
\textsc{Merge}, \textsc{Insert}, and \textsc{DeleteMax};
all can be made to have $O(\log n)$ worst-case behaviour, amortised.

Your task is to reconstruct pseudo-code for the three operations
(including some helper functions).
The appendix contains a scrambled solution,
with lines of pseudo-code in random order, without any indentation.
You are asked to reconstruct the original code, with proper indentation.

The data structure we use is a kind of \emph{multiway tree},
that is, a node can have any number of subtrees.
Since it is a multiway tree and also a (max-)\emph{heap},
we will call it a \emph{meap}.
A meap $M$ can be empty, in which case we denote it \textbf{void}.
If $M$ is non-empty then $M$ has a key, plus a list of non-empty
sub-trees (sub-meaps).
Note that the list of subtrees can be of any length;
in particular it can be empty.

A \emph{tree list} $L$ can be empty, in which case we denote it
\textbf{null}.
If $L$ is non-empty, then $L$ has a first tree, $L.\id{elt}$ and
a tail, $L.\id{next}$, with the remaining trees.
In our case, these trees are all meaps.

A meap $M$ (other than \textbf{void}) has two attributes:
\begin{itemize}
\setlength{\itemsep}{-0.5ex}
\item $M.\id{key}$ is the root node's key.
\item $M.\id{subtrees}$ is a tree list---the list of $M$'s sub-meaps.
\end{itemize}
A meap must satisfy the \emph{heap condition}:
The key of a child node is smaller than the key of its parent.
That is, for every meap $M'$ in $M.\id{subtrees}$,
we have $M'.\id{key} \leq M.\id{key}$.
The heap condition must hold recursively, for all nodes.

Below are two meaps, $M_1$ (on the left) and $M_2$ (on the right):
\begin{center}
\pstree[nodesep=2pt,levelsep=25pt,treesep=4mm]{\TR{92}}{
  \pstree[treesep=2mm]{\TR{75}}{
    \TR{16}
    \TR{52}
    \TR{27}
  }
  \TR{33}
  \pstree{\TR{66}}{
    \TR{44}
  }
  \pstree{\TR{49}}{
    \TR{7}
  }
  \TR{10}
}
\qquad \qquad
\pstree[nodesep=2pt,levelsep=25pt,treesep=4mm]{\TR{81}}{
  \TR{30}
  \pstree{\TR{64}}{
    \pstree{\TR{57}}{
      \TR{23}
      \TR{44}
    }
  }
}
\end{center}

\vspace{1ex}\noindent
Here $M_1.\id{key} = 92$ and $M_1.\id{subtrees}$ is a list of five children.
These children are meaps with roots 75, 33, 66, 49, and 10.

To find and return the largest key in a meap is easy---it is just the root's
key; we will not be worried about that operation in this question.

To merge two meaps, we include the meap that has the smaller root key
as the leftmost subtree of the other meap.
Merging the two meaps above, we get $\textsc{Merge}(M_1,M_2)$,
shown here:
\begin{center}
\pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{92}}{
  \pstree{\TR{81}}{
    \TR{30}
    \pstree{\TR{64}}{
      \pstree{\TR{57}}{
        \TR{23}
        \TR{44}
      }
    }
  }
  \pstree[treesep=2mm]{\TR{75}}{
    \TR{16}
    \TR{52}
    \TR{27}
  }
  \TR{33}
  \pstree{\TR{66}}{
    \TR{44}
  }
  \pstree{\TR{49}}{
    \TR{7}
  }
  \TR{10}
}
\end{center}

\vspace{1ex}\noindent
Once \textsc{Merge} has been implemented, \textsc{Insert} is easy.
\textsc{DeleteMax} removes the largest element from a meap.
It also makes use of \textsc{Merge} but is more complicated.
After discarding the root it must merge the children, to return a
single meap.
We do the merging in two passes.
First we merge the children in pairs from left to right
(that is, the first child with the second, the third with the fourth,
and so on).
Then we merge the resulting meaps in a single sweep from right to left.

Continuing our example, removal of the root (92) leaves us with
six meaps to merge:
\begin{center}
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{81}}{
    \TR{30}
    \pstree{\TR{64}}{
      \pstree{\TR{57}}{
        \TR{23}
        \TR{44}
      }
    }
  }
  \quad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{75}}{
    \TR{16}
    \TR{52}
    \TR{27}
  }
  \quad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{33}}{}
  \qquad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{66}}{
    \TR{44}
  }
  \qquad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{49}}{
    \TR{7}
  }
  \qquad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{10}}{}
\end{center}

\vspace{1ex}\noindent
Merging pairwise, from left to right, we get:
\begin{center}
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{81}}{
    \pstree[treesep=2mm]{\TR{75}}{
      \TR{16}
      \TR{52}
      \TR{27}
    }
    \TR{30}
    \pstree{\TR{64}}{
      \pstree{\TR{57}}{
        \TR{23}
        \TR{44}
      }
    }
  }
  \quad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{66}}{
    \TR{33}
    \TR{44}
  }
  \qquad
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{49}}{
    \TR{10}
    \TR{7}
  }
\end{center}

\pagebreak
\noindent
Now, merging the resulting meaps in a single pass from right to left,
we get:
\begin{center}
  \pstree[nodesep=2pt,levelsep=25pt,treesep=5mm]{\TR{81}}{
    \pstree[treesep=2mm]{\TR{66}}{
      \pstree{\TR{49}}{
        \TR{10}
        \TR{7}
      }
      \TR{33}
      \TR{44}
    }
    \pstree[treesep=2mm]{\TR{75}}{
      \TR{16}
      \TR{52}
      \TR{27}
    }
    \TR{30}
    \pstree{\TR{64}}{
      \pstree{\TR{57}}{
        \TR{23}
        \TR{44}
      }
    }
  }
\end{center}

\vspace{1ex}\noindent
Now have fun re-assembling the code from the appendix.
\hfill
\\
\textsc{answer}
\par
\begin{algorithmic}
  \Function{AddToList}{$M, L$}
    \State $R \gets \mathbf{new}\ \id{TreeList}$
    \State $R.\id{elt} \gets M$
    \State $R.\id{next} \gets L$
    \State \Return $R$
  \EndFunction

  \Function{Merge}{$M_1,M_2$}
    \If{$M_1 = \mathbf{void}$}
      \State \Return $M_2$
    \EndIf
    \If{$M_2 = \mathbf{void}$}
      \State \Return $M_1$
    \EndIf
    \State $M \gets \mathbf{new}\ \id{MultiwayTree}$
    \If{$M_1.\id{key} > M_2.\id{key}$}
      \State $M.\id{key} \gets M_1.\id{key}$
      \State $M.\id{subtrees} \gets \textsc{AddToList}(M_2,M_1.\id{subtrees})$
    \Else
      \State $M.\id{key} \gets M_2.\id{key}$
      \State $M.\id{subtrees} \gets \textsc{AddToList}(M_1,M_2.\id{subtrees})$
    \EndIf
    \State \Return $M$
  \EndFunction

  \Function{Insert}{$x, M$}
    \State $M' \gets \mathbf{new}\ \id{MultiwayTree}$
    \State $M'.\id{key} \gets x$
    \State $M'.\id{subtrees} \gets \mathbf{null}$
    \State \Return $\Call{Merge}{M',M}$
  \EndFunction

  \Function{MergePairs}{$L$}
    \If{$L = \mathbf{null}$}
      \State \Return $\mathbf{void}$
    \EndIf
    \State $M_1 \gets L.\id{elt}$
    \State $L_1 \gets L.\id{next}$
    \If{$L_1 = \mathbf{null}$}
      \State \Return $M_1$
    \EndIf
    \State $M_2 \gets L_1.\id{elt}$
    \State $L_2 \gets L_1.\id{next}$

    \State \Return $\Call{Merge}{{\Call{Merge}{M_1, M_2}},\Call{MergePairs}{L_2}}$
  \EndFunction

  \Function{DeleteMax}{$M$}
    \If{$M = \mathbf{void}$}
      \State \textsc{Error}(``Cannot delete from empty tree'')
    \EndIf
    \State \Return $\Call{MergePairs}{M.\id{subtrees}}$
  \EndFunction
\end{algorithmic}

\begin{flushright}
Tiancong Li
\\ 27 September 2017
\end{flushright}
\hfill

\section*{Appendix: Scrambled solution for Question 5}
\begin{algorithmic}
\State \textbf{else}
\State \textsc{Error}(``Cannot delete from empty tree'')
\State \textbf{function} $\textsc{AddToList}(M,L)$ : $\id{TreeList}$
\State \textbf{function} $\textsc{DeleteMax}(M)$ : $\id{MultiwayTree}$
\State \textbf{function} $\textsc{Insert}(x,M)$ : $\id{MultiwayTree}$
\State \textbf{function} $\textsc{Merge}(M_1,M_2)$ : $\id{MultiwayTree}$
\State \textbf{function} $\textsc{MergePairs}(L)$ : $\id{MultiwayTree}$
\State \textbf{if} $L = \mathbf{null}$ \textbf{then}
\State \textbf{if} $L_1 = \mathbf{null}$ \textbf{then}
\State \textbf{if} $M = \mathbf{void}$ \textbf{then}
\State \textbf{if} $M_1 = \mathbf{void}$ \textbf{then}
\State \textbf{if} $M_1.\id{key} > M_2.\id{key}$ \textbf{then}
\State \textbf{if} $M_2 = \mathbf{void}$ \textbf{then}
\State $L_1 \gets L.\id{next}$
\State $L_2 \gets L_1.\id{next}$
\State $M \gets \mathbf{new}\ \id{MultiwayTree}$
\State $M.\id{key} \gets M_1.\id{key}$
\State $M.\id{key} \gets M_2.\id{key}$
\State $M.\id{subtrees} \gets \textsc{AddToList}(M_1,M_2.\id{subtrees})$
\State $M.\id{subtrees} \gets \textsc{AddToList}(M_2,M_1.\id{subtrees})$
\State $M' \gets \mathbf{new}\ \id{MultiwayTree}$
\State $M'.\id{key} \gets x$
\State $M'.\id{subtrees} \gets \mathbf{null}$
\State $M_1 \gets L.\id{elt}$
\State $M_2 \gets L_1.\id{elt}$
\State $R \gets \mathbf{new}\ \id{TreeList}$
\State $R.\id{elt} \gets M$
\State $R.\id{next} \gets L$
\State \textbf{return} $M$
\State \textbf{return} $M_1$
\State \textbf{return} $M_1$
\State \textbf{return} $M_2$
\State \textbf{return} $\textsc{Merge}(M',M)$
\State \textbf{return}
       $\textsc{Merge}(\textsc{Merge}(M_1,M_2),\textsc{MergePairs}(L_2))$
\State \textbf{return} $\textsc{MergePairs}(M.\id{subtrees})$
\State \textbf{return} $R$
\State \textbf{return} $\mathbf{void}$
\end{algorithmic}

\end{document}
